The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2015-11-13 DishBeamMeasurements/paper/DishBeamMeasurementsPaper.tex

In detail, we discuss the experimental setup of the beam mapping experiments and their systematics, then 
review  the ORBCOMM beam measurement system, in Section 2. We discuss the electromagnetic design and 
modeling of the dish in Section 3, present our power pattern measurements in Section 5. We conclude with 
discussion in Section 6.

@@

\subsection{Design of the HERA Dish}

ASK DAVE DEBOER TO WRITE A TWO PARAGRAPH SUMMARY OF THE HERA ELEMENT AND HOW IT WAS DESIGNED

%\section{Feed Measurements}

%Accurate modeling of both the dish and feed will be essential for final dish modeling. To that end, we deploy the feed on the ground facing the sky (Figure \ref{fig:feedpic}), replacing the north-most reference antenna in the \texttt{null3} configuration, to isolate the feed power pattern.

%Figure \ref{fig:feed0} shows the measured feed beams (top panel) and slices through the E and H planes (bottom panel). Due to intermittent connection problems to the DAQ computer, we collected only 157 high signal-to-background satellite passes over a few days. Still, the data agree very well with the CST model for $\theta<55^\circ$, outside of which the E plane response starts narrows compared to the model by up to 5\,dB at $\theta=70^\circ$. Still, the data are quite sparse in these low regions due to the connection problems, and so the usual quality control cuts are likely less effective than usual.

%\begin{figure*}[h]
%\includegraphics[width=5in]{feed.jpg}
%\caption{Feed deployed on the ground, facing the sky, replacing the north-most reference antenna in the \texttt{null3} configuration. }
%\label{fig:feedpic}
%\end{figure*}

%\begin{figure*}[h]
%\includegraphics[width=6.5in]{feed0_abs_old_ref_model.png}
%\caption{Measurements of the feed power pattern, deployed so it faces the sky. The measurements agree well with the model for $\theta<55^\circ$, but start to diverge at large zenith angles, possibly due to quality control problems due to very sparse beam sampling.}
%\label{fig:feed0}
%\end{figure*}

>>>>

In detail, we discuss the electromagnetic design and modeling of the dish in Section 3. We present the 
experimental setup of the beam mapping experiments and discuss their systematics, then 
review the ORBCOMM beam measurement system, in Section 3. We present our power pattern 
measurements in Section 5. We conclude with discussion in Section 6.

\subsection{Design of the HERA Dish}
%- EM design
%    - departure from large N—small D to economize data volume (mitigate
%      data processing challenges)
%        - easier to identify systematics, to rerun datasets, perform
%          cross checks / jackknives
%    - but this change must be done judiciously, one advantage of simple
%      elements is that they have little or no frequency structure,
%      while a dish+feed system introduces reflections in the time
%      domain, amounting in ripples in the frequency response
%    - how bad are these ripples and can we tune the dish to mitigate
%      their effect in the science?
%    - as discussed above, it is useful to think in the delay domain as
%      the translation from frequency structure introduced by dish
%      reflections and the foreground isolation within the horizon delay
%      is straightforward
%    - the smooth spectral structure of the foregrounds sets a maximum
%      foreground isolation achievable of ~60ns with blackman-harris
%      window
%    - f/D =0.32 maximizes efficiency for a PAPER dipole suspended over
%      the dish
%    - with typical reflection dish/feed reflection coefs, attenuation
%      after 3 round trips will be below 60dB, and 1 and 2 round trips
%      are in wedge, isolated away from EOR window
%    - 14m also falls near the minimum of the cost per performance model

The 14\,m HERA dish design is a departure from the large N--small D approach used by 21\,cm 
observatories like PAPER and the MWA. Both observatories are actively pursuing power 
spectrum analyses using several year data runs, but the sheer data volume makes 
characterization and removal of systematics, as well as repeated or complimentary analyses, 
challenging. A larger element was chosen for HERA primarily to economize data volume. 
The natural consequence of a larger antenna aperture is a smaller field of view, but this is a 
small effect for 21\,cm power spectrum analyses as our leverage on $k$ modes in the 
spherically averaged power spectrum comes primarily from $k_\parallel$ modes along the line 
of sight (in the frequency dimension).

However, such a turn from small, simpler antenna elements to larger, more complex dishes with 
suspended feeds must be done judiciously, lest the chromatic antenna response smear otherwise 
smooth spectrum foreground power into cosmological signal modes. For this reason the dish 
design was carefully optimized for a foreground avoidance-based power spectrum spectrum 
analysis, as detailed by \citet{heraMemo5} and discussed in the larger engineering context of 
HERA by DeBoer et al (in prep).

We summarize here the logic leading to a 14\,m dish with $f/D\approx0.32$. As discussed 
above, it is useful to think in delay-space, where a source ideally appears as a delta function 
at the delay corresponding to the difference in light travel time to two antennas in a 
baseline. In a real instrument, finite bandwidth and intrinsic source spectral structure 
(synchrotron sources typically have $I\propto f^{-0.85}$) smear foreground power over a 
kernel as wide as $\sim60$\,ns. All antennas have some frequency structure as well, though 
dipole-like elements such as sleeved dipoles and bowties may be made relative frequency 
independent over wide bands. Suspension of a feed over a dish introduces frequency 
structure directly due to time domain reflections between the dish and feed. We optimize the 
HERA dish so this frequency structure extends no farther into delay space than the 60\,ns 
intrinsic width of foregrounds. This was accomplished first through numerical modeling of 
the beam pattern with a PAPER dipole suspended over the dish aperture, and resulted in a 
maximum efficiency of 73\% with $f/d\approx0.32$. From there, using realistic estimates of 
reflection loss, a 14\,m dish was observed to have the property that the signal received at the 
feed after two bounces (round trips between the dish and feed) is above -60\,dB but 
contained in the wedge, but all higher order reflections in the EOR window are below 
-60\,dB, and thus an order of magnitude smaller than the expected cosmological signal.

In reality, the HERA dishes are somewhat more complicated that these simple considerations 
imply, and part of the purpose of the beam measurements presented in this paper is to 
characterize their \textit{in situ} beam patterns. The dish is fabricated out of low cost 
construction materials with an expected lifetime of $\sim5$ years, sufficient to detect the 21\,cm 
power spectrum to high significance before the Square Kilometer Array builds a longer lasting 
instrument to more fully characterize the Epoch of Reionization with imaging. The feed consists 
of a sleeved dipole, suspended in a cage structure from from three ropes attaching to telephone 
poles spaced around the dish. The dipole is mounted on a XX\,m mast over a XX\,m diameter backplane surrounded by a XX\,m long cylindrical ``skirt'' designed to reduce cross coupling with adjacent elements 
by narrowing the feed's response. The dish surface consists of 12 wire mesh strips secured to 
60\,mm PVC pipes forming the dish skeleton. A spar at 2.4\,m elevation is added to ensure 
approximately parabolic shape. Note that we refer to this design as a \textit{faceted parabola} 
given that the dish surface between PVC spars is not curved, though our design ensures the 
Ruze loss is at the percent level. A door is engineered into one of these mesh strips to facilitate 
feed maintenance.

This work presents beam pattern measurements to verify the capabilities of this new dish, in particular the 
position of the dish focus, the shape of the main lobe, the magnitude of the dish sidelobes, the 
degree of beam symmetry, the actual dish/feed efficiency (i.e., collecting area), and the expected level of antenna-to-antenna variation.

@@
